The design and implementation of standards-based Grid single sign-on using federated identity

### 2009-05-01 doc/sec/papers/GRID09/paper.tex
More text

\begin{figure}
\includegraphics[width=0.9\columnwidth]{SAML2SSO_thread10_nosso.pdf}
\caption{Performance of HED echoString services with SAMLSSO profile: 1 service
and 10 clients}
\label{fig:SAML2SSO_thread10_nosso}
\end{figure}

\begin{figure}
\includegraphics[width=0.9\columnwidth]{SAML2SSO_thread10_nosso_and_sso.pdf}
\caption{Performance of HED echoString services with SAMLSSO profile: 1 service
and 10 clients}
\label{fig:SAML2SSO_thread10_nosso_and_sso}
\end{figure}

\begin{figure}
\includegraphics[width=0.9\columnwidth]{SAML2SSO_thread_all.pdf}
\caption{Performance of HED echoString services with SAMLSSO profile:
performance changes with client numbers; session non-reused }
\label{fig:SAML2SSO_thread_all}
\end{figure}

Figure \ref{fig:SAML2SSO_thread10_nosso} depicts the average response time and
throughputs for 10 clients running simultaneously against the ``echoString'' service enhanced with SAML2SSO
profile. Since the success of authentication from IdP can be regarded as valid for the whole
session between client and service, the client with ``Session reused'' need to accomplish the SAML2SSO
profile only once for all SOAP invocations on this client, while the client with ``Session
non-reused'' will need to accomplish this profile once per SOAP invocation.

The average response time of ``Session non-reused'' experiments changes little
with the execution duration, and the value is about 3000ms; while the throughput is around 200
requests per minutes. Considering that the average response time in case of ``Session non-reused'' and
pure SOAP invocation is around 15ms when client number is 10 (see section \ref{sec:perhedandaxis}),
therefore we can calculate that the consumption of time for SAML2SSO profile is about 99.5\% of
the whole SOAP invocation with SAML2SSO profile integrated.

@@

One the other hand, the average response time of ``Session reused'' experiments
is about 82 ms, which is almost as short as the value in the pure SOAP invocation experiment in
section \ref{sec:perhedandaxis}. The result is reasonable because the time consumption
caused by only once of SAML2SSO execution on one client is relatively small comparing with
the time consumption caused by the following hundreds of pure SOAP invocations on the
same client.

Figure~\ref{fig:SAML2SSO_thread_all} depicts the performance results for
echoString services that are enhanced with SAML2SSO profile, with client number being changed. In
this set of experiments, the session is not reused, and secure communication without client certificate
authentication is used. The results show the average response time increases linearly with the client
number, while the throughput increases initially and then keeps in a fixed value (around 215
requests per minute) when the client number is bigger than 4. When the client number is four, the
response time is around one second, while the throughput of the service is 215 requests per
minutes which is almost the fixed value.
We speculate that in order to get acceptable response time, there should be less
than four (including four) concurrent clients if they need to continuously participate one SAML2SSO
profile.

Considering that the throughput of the ``echoString'' service enhanced with mutual
TLS is 13000, and the concurrent client number when arriving fixed throughput is 30, the
performance of ``echoString'' service enhanced with SAML2SSO profile is much worse, with the
values being 215 and 4 correspondingly.

However for some typical grid application scenarios, normally the client does
not need to continuously accomplish SAML2SSO profile. For instance, a user could submit a
job which needs a period of time (normally this period is much longer than a few
seconds) to execute, which launches an execution of SAML2SSO profile; 
afterwards, this user could query the information as well as the result about
this job, which launches another execution of SAML2SSO profile. On the other hand,
statistically, users will most probably not access grid system in a concurrent way, rather than a random way.
Therefore, if we suppose a user could access grid once per minute, in our test environment, we
can claim that around 215 clients randomly can be supported with acceptable performance.

@@

\begin{figure}
\includegraphics[width=0.9\columnwidth]{SLCS_thread_all.pdf}
\caption{Performance of SLCS service: performance changes with client numbers}
\label{fig:SLCS_thread_all}
\end{figure}

@@

@@

\section*{Acknowledgment}

The authors would like to thank...

>>>>

\begin{figure}
\includegraphics[width=0.9\columnwidth]{SAML2SSO_thread10_nosso.pdf}
\caption{Performance of HED echoString services with SAML2SSO profile: 1 service
and 10 clients; single sign-on is disabled; session used and session not re-used are compared}
\label{fig:SAML2SSO_thread10_nosso}
\end{figure}

\begin{figure}
\includegraphics[width=0.9\columnwidth]{SAML2SSO_thread10_nosso_and_sso.pdf}
\caption{Performance of HED echoString services with SAML2SSO profile: 1 service
and 10 clients; session not re-used; single sign-on enabled and single sign-on disabled are compared}
\label{fig:SAML2SSO_thread10_nosso_and_sso}
\end{figure}

\begin{figure}
\includegraphics[width=0.9\columnwidth]{SAML2SSO_thread_all.pdf}
\caption{Performance of HED echoString services with SAML2SSO profile:
performance changes with client numbers; session not re-used; single sign-on enabled and 
single sign-on disabled are compared}
\label{fig:SAML2SSO_thread_all}
\end{figure}

We consider two situations in terms of the authenticating between user agent (client) and IdP: 
SSO enabled, and SSO disabled. ``SSO disabled'' means the successful authentication
session between user agent and IdP is not regarded as valid for the subsequent authentication, and
user is required to authenticate by inputing his username and password on each request.
``SSO enabled'' means the successful authentication session is regarded as valid for the 
subsequent authentication, in a short period (we setup this period as 30 minutes).

Considering the success of authentication from IdP can be regarded as valid for the whole
session between client and service, we also consider two situation in terms of the 
using of authentication result between client and service: session re-used, and session non-reused.
The client with ``Session reused'' need to accomplish the SAML2SSO
profile only once for all SOAP invocations on this client, while the client with ``Session
non-reused'' will need to accomplish this profile once per SOAP invocation.

Figure \ref{fig:SAML2SSO_thread10_nosso} depicts the average response time and
throughputs for 10 clients running simultaneously against the ``echoString'' service enhanced with SAML2SSO
profile, with single sign-on disabled.

The average response time of ``Session non-reused'' experiments changes little
with the execution duration, and the value is about 3000ms; while the throughput is around 215
requests per minutes. Considering that the average response time in case of ``Session non-reused'' and
pure SOAP invocation is around 15ms when client number is 10 (see section \ref{sec:perhedandaxis}),
therefore we can calculate that the consumption of time for SAML2SSO profile is about 99.5\% of
the whole SOAP invocation with SAML2SSO profile integrated.

@@

One the other hand, the average response time of ``Session reused'' experiments
is about 82 ms, which is almost as short as the value in the pure SOAP invocation experiment in
section \ref{sec:perhedandaxis}. The result is reasonable because the time consumption
caused by only once of SAML2SSO execution on one client is relatively small comparing with
the time consumption caused by the following hundreds of pure SOAP invocations on the
same client. So besides improving the performance of SAML2SSO profile, reusing 
the session as longer as possbile is also a way to improve the performance of 
Web Service invokation with SAML2SSO profile integrated as authentication approach.

Figure \ref{fig:SAML2SSO_thread10_nosso_and_sso} depicts the performance comparison between singled 
sign-on disabled and singled sign-on enabled, for 10 clients running simultaneously 
against the ``echoString'' service enhanced with SAML2SSO profile, with session not re-used.

The average response time of ``SSO enabled'' is about 1700ms, while the throughput is about
340 requests per minutes. The average response time of ``SSO disabled'' is about 3000ms, while 
the through put is about 215 requests per minutes. The comparison shows that the performance 
of ``SSO enabled'' is better than that of ``SSO disabled'', but still on the same quantitative
level.

Figure~\ref{fig:SAML2SSO_thread_all} depicts the performance results for
echoString services that are enhanced with SAML2SSO profile, with client number being changed. In
this set of experiments, the session is not reused, and single sign-on enabled and singled sign-on
disabled are compared.

For the situation of singled sign-on disabled, the average response time 
increases linearly with the client number, while the throughput increases initially and then 
keeps in a fixed value (around 215 requests per minute) when the client number is bigger 
than 4. When the client number is four, the response time is around one second, while the 
throughput of the service is 215 requests per minutes which is almost the fixed value.

For the situation of singled sign-on enabled, the average response time 
also increases linearly with the client number, while the throughput also increases initially and then 
keeps in a fixed value (around 340 requests per minute) when the client number is bigger 
than 5. When the client number is five, the response time is around 880ms, while the 
throughput of the service is 340 requests per minutes which is almost the fixed value.

We speculate that in order to get acceptable response time, for the case of single sign-on disabled, 
there should be less than four (including four) concurrent clients if they need to continuously 
participate one SAML2SSO profile; for the case of single sign-on enabled, 
there should be less than five (including five) concurrent clients if they need to continuously 
participate one SAML2SSO profile.

Considering that the throughput of the ``echoString'' service enhanced with mutual
TLS is 13000, and the concurrent client number when arriving fixed throughput is 30, the
performance of ``echoString'' service enhanced with SAML2SSO profile is much worse, both 
with single sign-on disabled and enabled, with the values being 215, 4, and 340, 5, correspondingly.

However for some typical grid application scenarios, normally the client does
not need to continuously accomplish SAML2SSO profile. For instance, a user could submit a
job which needs a period of time (normally this period is much longer than a few
seconds) to execute, which launches an execution of SAML2SSO profile; 
afterwards, this user could query the information as well as the result about
this job, which launches another execution of SAML2SSO profile. On the other hand,
statistically, users will most probably not access grid system in a concurrent way, rather than a random way.
Therefore, if we suppose a user could access grid once per minute, in our test environment, we
can claim that for the case of of single sign-on disabled, around 215 clients randomly can be 
supported with acceptable performance, while for the case of single sign-on enabled, around 340 
clients randomly can be supported with acceptable performance.

@@

\begin{figure}
\includegraphics[width=0.9\columnwidth]{SLCS_thread_all.pdf}
\caption{Performance of SLCS service: performance changes with client numbers; single sign-on disabled}
\label{fig:SLCS_thread_all}
\end{figure}

@@

Considering the time interval of the usage of SLCS client should be similar to the lifetime of SLCS
credential which is 12 hours by default, and the lifetime of successful authentication session is 
30 minutes by default, enabling single sign-on is not useful in real deployment because the successful 
authentication session will be expired before the next time the SLCS client is used, therefore, 
we do not enable single sign-on when testing the SLCS service.

@@

\section*{Acknowledgment}
We would thank all of the ARC middleware developers for
their intelligent work. We wish to thank Oxana Smirnova for
vital comments and proof-reading.
